Mixed language processing increases cross-language phonetic transfer in Bengali-English Bilinguals

### 2021-05-04 draft_glossa/glossa-template.tex
fixed tables, word count

\title[Mixed language processing increases cross-language phonetic transfer in Bengali-English Bilinguals}
% Optional short title inside square brackets, for the running headers.

\author[]% short form of the author names for the running header. If no short author is given, no authors print in the headers.
{%as many authors as you like, each separated by \AND.
  \spauthor{.\\ 
  \institute{.\\
  \small{%105, Bd. Raspail, 75005 Paris\\
	}
  }
  \AND
  \spauthor{.\\
  \institute{.\\
  \small{%Warmoesberg 26, 1000 Brussel\\
  }
  }%
}

@@

Word count: 4269

@@

%Studies which compare bilinguals speakers' phonological systems with monolinguals see them as relatively stable over time. The difference from monolingual norms is interpreted as the cumulative result of cross-language influence over long periods, \cite<e.g.>{guion2003vowel, caramazza1973acquisition}. However, longitudinal studies %(citations) 
%and between-subject comparisons of bilingual speakers who differ in the duration of L2 exposure suggest that interaction between the sound systems is dynamic. Over time, increasing exposure to an L2 can reduce effects of L1 influence, leading to a more fine-grained separation and native-like production of foreign contrasts-- e.g. \cite{bohn1992production} 
%found that an approximately 7 years' difference in L2 exposure between groups having the same age of acquisition led to a significant difference in the amount of cross-language transfer. These have been interpreted as changes to long-term phonological representations.

%Moreover, changes due to transfer are not unidirectional or irreversible. \cite{sancier1997gestural} first demonstrated that spending 2-5 months in an L1 or L2 environment causes productions in both languages to ``drift" towards the ambient language. This not only showed that cross-language interaction can be triggered in an order of months, but also that the effects can be reversed within a similar time range. A recent study by \cite{tobin2017phonetic} found comparable effects in an even shorter duration (2-4 weeks). A short-term longitudinal study by \cite{chang2012rapid} showed that over the course of the first five weeks of learning an L2, there was a gradual convergence of L1 towards L2. For some sounds and features (though not others, cf sec.\ref{asymmetry between sounds}), transfer was additive over time. How long these effects last in the absence of regular L2 input was not tested. 
% if effects are cumulative, then the difference between switched and non-switched tokes should be lesser in later test blocks (previous two paragraphs needed?)

@@

10 Bengali-English bilingual speakers (male=5, age range 19 to 28), completed the tasks. All the participants speak other languages in addition to Bengali and English. At the time of recording, they were living in or around the EFL University campus in Hyderabad, which has a linguistically diverse population. Speakers were selected on the basis of responses to a Language Background Questionnaire (sec.\ref{appendix}) to ensure comparable LSRW skills in L1 and L2 across participants, and were compensated for their time.

@@

\begin{table}[h]
	\caption{Fixed factors}
	\sffamily
	\centering
	\begin{tabular}{l l}
		\hline 
		Factor & Levels\\ 
		\hline
		Gender & male, female\\
		Vowel & \nt{2}, \nt{\ae} \\
		Task (corresponding to the two paradigms) & p (Picture: cued picture-naming, s (Sentence: code-switching)\\
		Context (corresponding to the experimental conditions) & e (English: unilingual), b (Bengali: mixed)\\ [1ex]
		\hline	
	\end{tabular}
	\label{table variables}
\end{table}

@@

\begin{table}[htbp] \centering 
	\caption{Model outputs for F1 at each of the five points in the vowel} 
	\label{table_f1_fixed_effects} 
	\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
		\\[-1.8ex]\hline 
		\hline \\[-1.8ex] 
		& \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
		\cline{2-6} 
		\\[-1.8ex] & F1\_5 & F1\_15 & F1\_25 & F1\_35 & F1\_45 \\ 
		\\[-1.8ex] & (Model 1) & (Model 2) & (Model 3) & (Model 4) & (Model 5)\\ 
		\hline \\[-1.8ex] 
		(Intercept) & 0.447$^{**}$ & 0.718$^{***}$ & 0.750$^{***}$ & 0.863$^{***}$ & 0.921$^{***}$ \\ 
		& (0.211) & (0.176) & (0.153) & (0.152) & (0.150) \\ 
		& & & & & \\ 
		Gendermale & $-$1.171$^{***}$ & $-$1.526$^{***}$ & $-$1.586$^{***}$ & $-$1.668$^{***}$ & $-$1.681$^{***}$ \\ 
		& (0.237) & (0.184) & (0.168) & (0.177) & (0.175) \\ 
		& & & & & \\ 
		Contexte & $-$0.010 & $-$0.015 & 0.033 & 0.025 & 0.037 \\ 
		& (0.058) & (0.057) & (0.042) & (0.040) & (0.035) \\ 
		& & & & & \\ 
		Vowel\textbackslash vt & 0.022 & $-$0.070 & $-$0.034 & $-$0.131 & $-$0.216 \\ 
		& (0.170) & (0.175) & (0.170) & (0.174) & (0.168) \\ 
		& & & & & \\ 
		Tasks & 0.319$^{***}$ & 0.235$^{***}$ & 0.172$^{**}$ & 0.133$^{*}$ & 0.102 \\ 
		& (0.091) & (0.083) & (0.081) & (0.077) & (0.073) \\ 
		& & & & & \\ 
		Contexte:Vowel\textbackslash vt & $-$0.126$^{**}$ & $-$0.138$^{**}$ & $-$0.185$^{***}$ & $-$0.191$^{***}$ & $-$0.192$^{***}$ \\ 
		& (0.056) & (0.060) & (0.055) & (0.064) & (0.066) \\ 
		& & & & & \\ 
		\hline \\[-1.8ex] 
		Observations & 2,333 & 2,333 & 2,333 & 2,333 & 2,333 \\ 
		Log Likelihood & $-$1,771.558 & $-$1,589.316 & $-$1,278.310 & $-$1,313.795 & $-$1,380.083 \\ 
		Akaike Inf. Crit. & 3,599.116 & 3,234.631 & 2,612.621 & 2,683.589 & 2,816.165 \\ 
		Bayesian Inf. Crit. & 3,760.253 & 3,395.769 & 2,773.758 & 2,844.727 & 2,977.303 \\ 
		\hline 
		\hline \\[-1.8ex] 
		\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
	\end{tabular} 
\end{table}

@@

\begin{table}[htbp] \centering 
	\caption{Model outputs for F2 at each of the five points in the vowel} 
	\label{table_F2_fixed_effects} 
	\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
		\\[-1.8ex]\hline 
		\hline \\[-1.8ex] 
		& \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
		\cline{2-6} 
		\\[-1.8ex] & F2\_5 & F2\_15 & F2\_25 & F2\_35 & F2\_45 \\ 
		\\[-1.8ex] & (Model 1) & (Model 2) & (Model 3) & (Model 4) & (Model 5)\\ 
		\hline \\[-1.8ex] 
		(Intercept) & 0.607$^{***}$ & 0.838$^{***}$ & 0.778$^{***}$ & 0.827$^{***}$ & 0.826$^{***}$ \\ 
		& (0.222) & (0.213) & (0.211) & (0.203) & (0.204) \\ 
		& & & & & \\ 
		Gendermale & $-$0.839$^{***}$ & $-$0.891$^{***}$ & $-$0.807$^{***}$ & $-$0.750$^{***}$ & $-$0.748$^{***}$ \\ 
		& (0.124) & (0.106) & (0.107) & (0.116) & (0.109) \\ 
		& & & & & \\ 
		Contexte & 0.003 & $-$0.088$^{**}$ & $-$0.065$^{**}$ & $-$0.063 & $-$0.103$^{***}$ \\ 
		& (0.036) & (0.036) & (0.032) & (0.044) & (0.033) \\ 
		& & & & & \\ 
		Vowel\textbackslash vt & $-$0.217 & $-$0.615$^{***}$ & $-$0.651$^{***}$ & $-$0.801$^{***}$ & $-$0.836$^{***}$ \\ 
		& (0.175) & (0.183) & (0.175) & (0.168) & (0.166) \\ 
		& & & & & \\
		Tasks & $-$0.234$^{***}$ & $-$0.194$^{***}$ & $-$0.152$^{***}$ & $-$0.141$^{***}$ & $-$0.092$^{**}$ \\ 
		& (0.029) & (0.035) & (0.050) & (0.047) & (0.045) \\ 
		& & & & & \\
		Contexte:Vowel\textbackslash vt & 0.031 & 0.110$^{***}$ & 0.128$^{***}$ & 0.091$^{**}$ & 0.147$^{***}$ \\ 
		& (0.050) & (0.041) & (0.040) & (0.046) & (0.043) \\ 
		& & & & & \\   
		\hline \\[-1.8ex] 
		Observations & 2,333 & 2,333 & 2,333 & 2,333 & 2,333 \\ 
		Log Likelihood & $-$1,283.531 & $-$1,478.494 & $-$1,174.087 & $-$1,222.682 & $-$1,019.033 \\ 
		Akaike Inf. Crit. & 2,623.062 & 3,012.988 & 2,404.175 & 2,501.363 & 2,094.066 \\ 
		Bayesian Inf. Crit. & 2,784.199 & 3,174.125 & 2,565.312 & 2,662.501 & 2,255.204 \\ 
		\hline 
		\hline \\[-1.8ex] 
		\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
	\end{tabular} 
\end{table}

@@

%\nocite{*} %this is to get all the entries of the sample bibliography; delete this line for an actual Glossa submission

%\printbibliography %for use with biblatex; comment out if you use natbib
\bibliography{references,sample} %for use with natbib; comment out if you use biblatex, and change 'sample' by the name of your bib-file

>>>>

\title{Mixed language processing increases cross-language phonetic transfer in Bengali-English Bilinguals}
% Optional short title inside square brackets, for the running headers.

\author[]% short form of the author names for the running header. If no short author is given, no authors print in the headers.
{%as many authors as you like, each separated by \AND.
  \spauthor{.\\ 
  \institute{.\\
  \small{.\\
	}
  }
  \AND
  \spauthor{.\\
  \institute{.\\
  \small{.\\
  }
  }%
}

@@

Word count: 6944

@@

Studies which compare bilinguals speakers' phonological systems with monolinguals see them as relatively stable over time. The difference from monolingual norms is interpreted as the cumulative result of cross-language influence over long periods, \cite<e.g.>{guion2003vowel, caramazza1973acquisition}. However, longitudinal studies %(citations) 
and between-subject comparisons of bilingual speakers who differ in the duration of L2 exposure suggest that interaction between the sound systems is dynamic. Over time, increasing exposure to an L2 can reduce effects of L1 influence, leading to a more fine-grained separation and native-like production of foreign contrasts-- e.g. \cite{bohn1992production} 
found that an approximately 7 years' difference in L2 exposure between groups having the same age of acquisition led to a significant difference in the amount of cross-language transfer. These have been interpreted as changes to long-term phonological representations.

Moreover, changes due to transfer are not unidirectional or irreversible. \cite{sancier1997gestural} first demonstrated that spending 2-5 months in an L1 or L2 environment causes productions in both languages to ``drift" towards the ambient language. This not only showed that cross-language interaction can be triggered in an order of months, but also that the effects can be reversed within a similar time range. A recent study by \cite{tobin2017phonetic} found comparable effects in an even shorter duration (2-4 weeks). A short-term longitudinal study by \cite{chang2012rapid} showed that over the course of the first five weeks of learning an L2, there was a gradual convergence of L1 towards L2. For some sounds and features (though not others, cf sec.\ref{asymmetry between sounds}), transfer was additive over time. How long these effects last in the absence of regular L2 input was not tested. 
% if effects are cumulative, then the difference between switched and non-switched tokes should be lesser in later test blocks (previous two paragraphs needed?)

@@

10 Bengali-English bilingual speakers (male=5, age range 19 to 28), completed the tasks. All the participants speak other languages in addition to Bengali and English. At the time of recording, they were living in or around the EFL University campus in Hyderabad, which has a linguistically diverse population. Speakers were selected on the basis of responses to a Language Background Questionnaire to ensure comparable LSRW skills in L1 and L2 across participants, and were compensated for their time.

@@

\begin{table}
	\caption{Fixed factors}
	\sffamily
	\centering
	\begin{tabular}{p{0.5\textwidth}p{0.5\textwidth}}
		\hline 
		Factor & Levels\\ 
		\hline
		Gender & male, female\\
		Vowel & \nt{2}, \nt{\ae} \\
		Task (corresponding to the two paradigms) & p (Picture: cued picture-naming, s (Sentence: code-switching)\\
		Context (corresponding to the experimental conditions) & e (English: unilingual), b (Bengali: mixed)\\ [1ex]
		\hline	
	\end{tabular}
	\label{table variables}
\end{table}

@@

\begin{table}[htbp] \centering 
	\caption{Model outputs for F1 at each of the five points in the vowel} 
	\label{table_f1_fixed_effects} 
	\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
		\\[-1.8ex]\hline 
		\hline \\[-1.8ex] 
		& \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
		\cline{2-6} 
		\\[-1.8ex] & F1\_5 & F1\_15 & F1\_25 & F1\_35 & F1\_45 \\ 
		\\[-1.8ex] & (Model 1) & (Model 2) & (Model 3) & (Model 4) & (Model 5)\\ 
		\hline \\[-1.8ex] 
		(Intercept) & 0.447$^{**}$ & 0.718$^{***}$ & 0.750$^{***}$ & 0.863$^{***}$ & 0.921$^{***}$ \\ 
		& (0.211) & (0.176) & (0.153) & (0.152) & (0.150) \\ 
		& & & & & \\ 
		Gendermale & $-$1.171$^{***}$ & $-$1.526$^{***}$ & $-$1.586$^{***}$ & $-$1.668$^{***}$ & $-$1.681$^{***}$ \\ 
		& (0.237) & (0.184) & (0.168) & (0.177) & (0.175) \\ 
		& & & & & \\ 
		Contexte & $-$0.010 & $-$0.015 & 0.033 & 0.025 & 0.037 \\ 
		& (0.058) & (0.057) & (0.042) & (0.040) & (0.035) \\ 
		& & & & & \\ 
		Vowel\nt{2} & 0.022 & $-$0.070 & $-$0.034 & $-$0.131 & $-$0.216 \\ 
		& (0.170) & (0.175) & (0.170) & (0.174) & (0.168) \\ 
		& & & & & \\ 
		Tasks & 0.319$^{***}$ & 0.235$^{***}$ & 0.172$^{**}$ & 0.133$^{*}$ & 0.102 \\ 
		& (0.091) & (0.083) & (0.081) & (0.077) & (0.073) \\ 
		& & & & & \\ 
		Contexte:Vowel\nt{2} & $-$0.126$^{**}$ & $-$0.138$^{**}$ & $-$0.185$^{***}$ & $-$0.191$^{***}$ & $-$0.192$^{***}$ \\ 
		& (0.056) & (0.060) & (0.055) & (0.064) & (0.066) \\ 
		& & & & & \\ 
		\hline \\[-1.8ex] 
		Observations & 2,333 & 2,333 & 2,333 & 2,333 & 2,333 \\ 
		Log Likelihood & $-$1,771.558 & $-$1,589.316 & $-$1,278.310 & $-$1,313.795 & $-$1,380.083 \\ 
		Akaike Inf. Crit. & 3,599.116 & 3,234.631 & 2,612.621 & 2,683.589 & 2,816.165 \\ 
		Bayesian Inf. Crit. & 3,760.253 & 3,395.769 & 2,773.758 & 2,844.727 & 2,977.303 \\ 
		\hline 
		\hline \\[-1.8ex] 
		\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
	\end{tabular} 
\end{table}

@@

\begin{table}[htbp] \centering 
	\caption{Model outputs for F2 at each of the five points in the vowel} 
	\label{table_F2_fixed_effects} 
	\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
		\\[-1.8ex]\hline 
		\hline \\[-1.8ex] 
		& \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
		\cline{2-6} 
		\\[-1.8ex] & F2\_5 & F2\_15 & F2\_25 & F2\_35 & F2\_45 \\ 
		\\[-1.8ex] & (Model 1) & (Model 2) & (Model 3) & (Model 4) & (Model 5)\\ 
		\hline \\[-1.8ex] 
		(Intercept) & 0.607$^{***}$ & 0.838$^{***}$ & 0.778$^{***}$ & 0.827$^{***}$ & 0.826$^{***}$ \\ 
		& (0.222) & (0.213) & (0.211) & (0.203) & (0.204) \\ 
		& & & & & \\ 
		Gendermale & $-$0.839$^{***}$ & $-$0.891$^{***}$ & $-$0.807$^{***}$ & $-$0.750$^{***}$ & $-$0.748$^{***}$ \\ 
		& (0.124) & (0.106) & (0.107) & (0.116) & (0.109) \\ 
		& & & & & \\ 
		Contexte & 0.003 & $-$0.088$^{**}$ & $-$0.065$^{**}$ & $-$0.063 & $-$0.103$^{***}$ \\ 
		& (0.036) & (0.036) & (0.032) & (0.044) & (0.033) \\ 
		& & & & & \\ 
		Vowel\nt{2} & $-$0.217 & $-$0.615$^{***}$ & $-$0.651$^{***}$ & $-$0.801$^{***}$ & $-$0.836$^{***}$ \\ 
		& (0.175) & (0.183) & (0.175) & (0.168) & (0.166) \\ 
		& & & & & \\
		Tasks & $-$0.234$^{***}$ & $-$0.194$^{***}$ & $-$0.152$^{***}$ & $-$0.141$^{***}$ & $-$0.092$^{**}$ \\ 
		& (0.029) & (0.035) & (0.050) & (0.047) & (0.045) \\ 
		& & & & & \\
		Contexte:Vowel\nt{2} & 0.031 & 0.110$^{***}$ & 0.128$^{***}$ & 0.091$^{**}$ & 0.147$^{***}$ \\ 
		& (0.050) & (0.041) & (0.040) & (0.046) & (0.043) \\ 
		& & & & & \\   
		\hline \\[-1.8ex] 
		Observations & 2,333 & 2,333 & 2,333 & 2,333 & 2,333 \\ 
		Log Likelihood & $-$1,283.531 & $-$1,478.494 & $-$1,174.087 & $-$1,222.682 & $-$1,019.033 \\ 
		Akaike Inf. Crit. & 2,623.062 & 3,012.988 & 2,404.175 & 2,501.363 & 2,094.066 \\ 
		Bayesian Inf. Crit. & 2,784.199 & 3,174.125 & 2,565.312 & 2,662.501 & 2,255.204 \\ 
		\hline 
		\hline \\[-1.8ex] 
		\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
	\end{tabular} 
\end{table}

@@

%\printbibliography %for use with biblatex; comment out if you use natbib
\bibliography{references} %for use with natbib; comment out if you use biblatex, and change 'sample' by the name of your bib-file
